---
title: "Reproducible Research: Peer Assessment 1"
author: "massuti"
date: "Sunday, February 15, 2015"
output: 
  html_document:
    keep_md: true
---  

## Loading and preprocessing the data  

```{r, echo=TRUE}
activity <- read.csv("activity.csv")
```  

## What is mean total number of steps taken per day?  

```{r, echo=TRUE}
stepbydate <- aggregate(steps ~ date, 
                        data=activity, sum, 
                        na.rm = TRUE)

print(stepbydate)

hist(stepbydate$steps, 
     main = "Steps by date", 
     breaks = 20)

mean(stepbydate$steps)

median(stepbydate$steps)
```  

## What is the average daily activity pattern?  

```{r, echo=TRUE}
stepsbyinterval <- aggregate(steps ~ interval, 
                             data=activity, mean, 
                             na.rm = TRUE)

plot(stepsbyinterval$interval, 
     stepsbyinterval$steps, 
     main = "Time Series of mean steps by intervals", 
     type = "l")
```  


Interval with maximum number of steps:
```{r, echo=TRUE}
stepsbyinterval[which.max(stepsbyinterval$steps), ]
```  

## Imputing missing values  


Total number of NA values:
```{r, echo=TRUE}
sum(is.na(activity))
```  


Filling NA values with mean steps in each 5 minute time interval and creating a new data set:
```{r, echo=TRUE}
activityNA <- data.frame(activity$steps)

activityNA[is.na(activityNA), ] <- floor(
        tapply(X=activity$steps, 
        INDEX=activity$interval,
        FUN=mean, na.rm = TRUE))

newactivity <- cbind(activityNA, activity[, 2:3])
colnames(newactivity) <- c("steps", "date", "interval")
```    
  
Making histogram of new data set steps by interval, to see changes:
```{r, echo=TRUE}
newstepbydate <- aggregate(steps ~ date, 
                        data=newactivity, sum)

print(newstepbydate)

hist(newstepbydate$steps,
     main = "Steps by date after filling NA",
     breaks = 20)

mean(newstepbydate$steps)

median(newstepbydate$steps)
```  

The NA values where substituted for mean values, bringing the mean values in the histogram to go up, as seen in the comparation of both,
```{r, echo=FALSE}
par(mfrow=c(1,2))

hist(newstepbydate$steps, 
     main = "ORIGINAL - Steps by date", 
     breaks = 20, ylim=c(0,18))

hist(stepbydate$steps, 
     main = "NEW - Steps by date", 
     breaks = 20, ylim=c(0,18))
```  

## Are there differences in activity patterns between weekdays and weekends?  

Creating a new variable:
```{r, echo=TRUE}
newactivity$day <- weekdays(as.Date(newactivity$date))

for (i in 1:nrow(newactivity)) {                                        
        if (newactivity[i,]$day %in% c("Saturday",
                                       "Sunday")) {
                newactivity[i,]$day <- "weekend"
                } else { newactivity[i,]$day <- "weekday"}
}

activitybyday <- aggregate(data=newactivity,
                           steps ~ day + interval,
                           FUN=mean)
```   

Ploting diference:
```{r, echo=TRUE}
library("lattice")

xyplot(type="l",
       data=activitybyday,
       steps ~ interval | day,
       layout=c(1,2))
```  
  
END.

